---
output:
  word_document: default
  pdf_document: default
  html_document: default
---
# Methods {#Methods}
```{r tidyr2, echo = FALSE, message = FALSE, warning = FALSE}

library(knitr)
library(pander)
library(suncalc)

opts_chunk$set(tidy.opts=list(width.cutoff=50), tidy = TRUE)
```
## Data Download
Least Tern detection data was downloaded from Motus using the motus R package, and then filtered to remove false positive detections. Detailed information on data download processes, formats, and false detections can be found here: [https://motus.org/MotusRBook/](https://motus.org/MotusRBook/). 

## Data Cleaning
Strings of detections with only 2 consecutive detections at a station (run length = 2) are questionable and may indicate the tag was just barely within range of the station, or more likely, that it is a false positive produced by background radio noise which is present to varying degrees at all stations. 

To begin, several tags were never detected near the breeding site and were removed (2017 Lotek tag ID's 95, 96, 97, 2018 Lotek tag ID 315). Then, detections at stations outside the tagging area with a run length of 2 were deemed to be false positives and removed. Each remaining tag was then visually inspected to examine detections that occurred at questionable latitudes far from the breeding site. Several stations had large numbers of detections with low run lengths, this is characteristic of stations with high levels of background radio noise which can often result in false positives. These stations were *Drasher*, *RTNJ*, *Koffler*, *Waggoners Gap*, *Burntpoint*, *Los Vientos Forest*, and *D'Estimauville*. Further inspection of detections at these stations showed that it would be biologically impossible for birds to make the flight between detections on the breeding ground and subsequent detections at these stations in the amount of time between consecutive detections. Therefore, all detections at these stations were deemed to be false positives due to noisy sites.

Remaining detections far from the tagging location were examined, many had run lengths of only 2 which are questionable at best and were deemed to be false positives. Others with run lengths of 3 or 4 were not biologically possible flights such as between the tagging area and eastern Canada. 

All detections deemed to be false positives were added to a custom filter that was applied to the data and were excluded from further analysis.

Detailed data filtering methodology with complete R code is available in [Appendix A](#AppendixA).

```{r LoadFilteredData, echo = FALSE, warning = FALSE, message = FALSE}
# load filtered data
library(motus)
proj.num <- 132
sql.motus <- tagme(projRecv = proj.num, new = FALSE, update = FALSE) # could check periodically if more data is expected 
tbl.alltags <- tbl(sql.motus, "alltags")

# obtain a table object of the filter
tbl.filter = getRunsFilters(sql.motus, "filtFalsePos")

# filter and convert the table into a data frame, with a few modifications
lete <- left_join(tbl.alltags, tbl.filter, by = c("runID", "motusTagID")) %>%
  mutate(probability = ifelse(is.na(probability), 1, probability),
         recvLat = if_else((is.na(gpsLat)|gpsLat == 0), # create new lat/lon variables
                           recvDeployLat, gpsLat), 
         recvLon = if_else((is.na(gpsLon)|gpsLon == 0), 
                           recvDeployLon, gpsLon)) %>% 
  filter(probability > 0) %>%
  #remove unnecessary columns
  select(-ambigID, -tagProjID, -sigsd, -noise, -freq, -freqsd, -slop, -burstSlop, -done, -bootnum, -tagType, -codeSet, -mfg, -tagModel, -tagLifespan, -nomFreq, -pulseLen, -markerNumber, -markerType, -tagDeployAlt, -tagDeployComments, -fullID, -recvDeployAlt, -antHeight, -speciesFR, -speciesSci, -speciesGroup, -tagProjName, -gpsAlt, -recvSiteName, -tagBI, -speciesID, -deviceID, -tagDeployID) %>%  
  filter(speciesEN == "Least Tern") %>% # keep only LETE
  collect() %>% 
  as.data.frame %>% # convert tbl to data.frame
  # convert times to datetime
  mutate(ts = as_datetime(ts, tz = "UTC", origin = "1970-01-01"),
         tagDeployStart = as_datetime(tagDeployStart, tz = "UTC", origin = "1970-01-01"),
         tagDeployEnd = as_datetime(tagDeployEnd, tz = "UTC", origin = "1970-01-01"),
         year = year(ts),
         recvDeployName = as.factor(recvDeployName),
         mfgID = as.numeric(mfgID)) %>% # get year variable
  arrange(motusTagID, ts)
# convert Lotek receivers to UTC time
lete$ts <- ifelse(lete$recvProjID %in% c(31, 132), (lete$ts + 5*60*60), lete$ts)
lete <- lete %>% mutate(ts = as_datetime(ts, tz = "UTC", origin = "1970-01-01"))

```

## Data Manipulation
Detailed nest monitoring records were used to determine tag retention and incubation periods. To determine incubation period, each tag was assigned an end date based on the "Fate" variable, once a nest's fate was categorized as either "failed nest", "successful nest", or "eggs due to hatch but no chicks found", it was assigned an end date of that day; the "nestEnd" variable.

Lotek tag ID's 114 and 316 did not have any nest records, as their incubation period could therefore not be determined, they were excluded from further analysis.

```{r importTagRetention, echo = FALSE, warning = FALSE, message = FALSE}
# Import nest monitoring data
nesting <- read.csv("./data/LETE Nesting & Tag Retention Data.csv")
# rename some variables
nesting <- rename(nesting, "mfgID" = "Tag..", "nestLat" = "Latitude", "nestLon" = "Longitude")
# convert date to proper date format
nesting <- mutate(nesting, Date = as.Date(nesting$Date, tz = "CDT"),
                  mfgID = as.character(mfgID))
# tags 132 and 120 are recorded together, need to separate
tag132 <- filter(nesting, mfgID == "132 & 120")
tag132$mfgID <- ifelse(tag132$mfgID == "132 & 120", "132", "")
nesting$mfgID <- ifelse(nesting$mfgID == "132 & 120", "120", nesting$mfgID)
nesting <- rbind(tag132, nesting)

# get data frame of nest end dates
nestEnd <- filter(nesting, is.na(Fate) == FALSE & mfgID != "")
nestEnd <- select(nestEnd, mfgID, Date)
nestEnd <- rename(nestEnd, "nestEnd" = "Date")
# get data frame of tag loss dates
tagLoss <- filter(nesting, Tag.Status == 2)
tagLoss <- tagLoss %>% group_by(mfgID) %>% summarize(tagLoss = min(Date)) %>% as.data.frame()
tagLoss$mfgID <- as.numeric(tagLoss$mfgID)
tags <- unique(select(nesting, mfgID))
tags$mfgID <- as.character(tags$mfgID)
tags$mfgID <- as.numeric(tags$mfgID)
tagLoss <- left_join(tags, tagLoss)
tagLoss$tagLoss[is.na(tagLoss$tagLoss)] <- as.Date("2019-01-01")
tagLoss <- tagLoss[!is.na(tagLoss$mfgID),]
# join nest end dates to nesting data frame
nesting <- left_join(select(nesting, mfgID, Nest_ID, nestLat, nestLon, Fate), nestEnd, by = "mfgID")
nesting <- unique(filter(nesting, mfgID != "", is.na(Fate) == FALSE))
nesting$mfgID <- as.numeric(nesting$mfgID)
# join tagLoss to nesting data frame
nesting <- left_join(nesting, tagLoss)
# join with lete database
lete <- left_join(lete, nesting, by = "mfgID")
# remove any detections after tag fell off
lete <- filter(lete, ts < tagLoss)
# label as incubating or not based on nestEnd date (ie. remove anything beyond nestEnd)
lete$incubate <- ifelse(as.Date(lete$ts) <= lete$nestEnd, "TRUE", "FALSE")
# Tags 114 and 316 do not have incubation information and will be set to FALSE
lete$incubate <- ifelse(lete$mfgID %in% c(114, 316), "FALSE", lete$incubate)

# Determine sunrise and sunset times for each detection based on detection location
lete <- sunRiseSet(lete, lat = "recvLat", lon = "recvLon")

#Only detections that occurred during incubation will be kept, 
# note that tags 114 and 316 do not have incubation information 
# and will thus be excluded.
inc <- filter(lete, incubate == TRUE)
```
Several tags were only detected for a short period of time; for meaningful results, tags that were detected for less than 24 hours were excluded from further analysis (Lotek tag ID's 85, 98, 120, 130, 286, 288, 309, 312). See [table 1.1](#tab:detectionlength) for a detection summary of all tags during incubation.
```{r detectionLength, echo = FALSE, message = FALSE, warning = FALSE}
#Examine how long each tag was detected for during incubation
detDate <- inc %>% group_by(year, motusTagID, mfgID) %>% 
  summarize(mints = min(ts), 
            maxts = max(ts), 
            range = round(as.numeric(difftime(maxts, mints, units = "days")), digits = 2),
            nSites = length(unique(recvDeployName))) %>% data.frame()
# some birds not detected for very long, therefore only look at ones detected for at least 1 full day
tmp <- unique(filter(detDate, range > 1)$motusTagID)
inc <- filter(inc, motusTagID %in% tmp)
tmp <- detDate
tmp <- rename(detDate, Year = year, "First Detection" = mints, "Final Detection" = maxts, "Detection Period (days)" = range, "Number of Sites" = nSites)
tmp <- tmp %>% arrange(motusTagID)
#kable(tmp, booktabs = TRUE, caption = 'Incubation detection period for all tags', longtable = TRUE)
pander(tmp, booktabs = TRUE, caption = '(\\#tab:detectionLength) Incubation detection period for all tags', split.cell = 11, split.table = Inf, justify = 'left')

```

```{r distances, echo = FALSE, message = FALSE, warning = FALSE}
#Create a data frame with distances between each receiver

# first create latLonDist function:
##
## translated from: http://www.movable-type.co.uk/scripts/latlong-vincenty.html
##
## /* - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -  */
## /* Vincenty Inverse Solution of Geodesics on the Ellipsoid (c) Chris Veness 2002-2012             */
## /*                                                                                                */
## /* from: Vincenty inverse formula - T Vincenty, "Direct and Inverse Solutions of Geodesics on the */
## /*       Ellipsoid with application of nested equations", Survey Review, vol XXII no 176, 1975    */
## /*       http://www.ngs.noaa.gov/PUBS_LIB/inverse.pdf                                             */
## /* - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -  */

## /**
##  * Calculates geodetic distance between two points specified by latitude/longitude using 
##  * Vincenty inverse formula for ellipsoids
##  *
##  * @param   {Number} lat1, lon1: first point in decimal degrees
##  * @param   {Number} lat2, lon2: second point in decimal degrees
##  * @returns (Number} distance in metres between points
##  */

latLonDist = function(lat1, lon1, lat2, lon2) {
  a = 6378137
  b = 6356752.314245
  f = 1/298.257223563  ## WGS-84 ellipsoid params
  
  llmat = cbind(lat1, lon1, lat2, lon2) ## recycles coordinates to match
  
  s = rep(-1, nrow(llmat)) ## return values; -1 means not yet computed
  for (i in 1:nrow(llmat)) {  ## calculate distance between i'th pair of points
    if (!all(is.finite(llmat[i,]))) {
      s[i] = NA
      next
    }
    
    L = rad(llmat[i, 4]-llmat[i, 2])
    U1 = atan((1-f) * tan(rad(llmat[i, 1])))
    U2 = atan((1-f) * tan(rad(llmat[i, 3])))
    sinU1 = sin(U1)
    cosU1 = cos(U1)
    sinU2 = sin(U2)
    cosU2 = cos(U2)
    lambda = L
    iterLimit = 100
    repeat {
      sinLambda = sin(lambda)
      cosLambda = cos(lambda)
      sinSigma = sqrt((cosU2*sinLambda) * (cosU2*sinLambda) + 
                        (cosU1*sinU2-sinU1*cosU2*cosLambda) * (cosU1*sinU2-sinU1*cosU2*cosLambda))
      if (abs(sinSigma) < 1e-10) {
        s[i] = 0 ## co-incident points
        break
      }
      cosSigma = sinU1*sinU2 + cosU1*cosU2*cosLambda
      sigma = atan2(sinSigma, cosSigma)
      sinAlpha = cosU1 * cosU2 * sinLambda / sinSigma
      cosSqAlpha = 1 - sinAlpha*sinAlpha
      cos2SigmaM = cosSigma - 2*sinU1*sinU2 / cosSqAlpha
      if (is.nan(cos2SigmaM))
        cos2SigmaM = 0  ## equatorial line: cosSqAlpha=0 (ยง6)
      C = f/16*cosSqAlpha*(4+f*(4-3*cosSqAlpha))
      lambdaP = lambda
      lambda = L + (1-C) * f * sinAlpha *
        (sigma + C*sinSigma*(cos2SigmaM+C*cosSigma*(-1+2*cos2SigmaM*cos2SigmaM)))
      iterLimit = iterLimit - 1
      if (abs(lambda-lambdaP) <= 1e-12 || iterLimit == 0)
        break
    } 
    
    if (iterLimit==0) {
      s[i] = NaN  ## formula failed to converge
    } else if (s[i] < 0) {
      uSq = cosSqAlpha * (a*a - b*b) / (b*b)
      A = 1 + uSq/16384*(4096+uSq*(-768+uSq*(320-175*uSq)))
      B = uSq/1024 * (256+uSq*(-128+uSq*(74-47*uSq)))
      deltaSigma = B*sinSigma*(cos2SigmaM+B/4*(cosSigma*(-1+2*cos2SigmaM*cos2SigmaM)-
                                                 B/6*cos2SigmaM*(-3+4*sinSigma*sinSigma)*(-3+4*cos2SigmaM*cos2SigmaM)))
      s[i] = b*A*(sigma-deltaSigma)
    }
  }
  s = round(s, 3)
  return (s)
}

## convert to radians from degrees
rad = function(x) {
  return (x * (pi/180))
}
## Distances
dist <- expand.grid("recvDeployName1" = unique(lete$recvDeployName), "recvDeployName2" = unique(lete$recvDeployName))
dist <- unique(left_join(dist, select(lete, recvDeployName, recvLat, recvLon), by = c("recvDeployName1" = "recvDeployName")))
dist <- rename(dist, "recvLat1" = "recvLat", "recvLon1" = "recvLon")
dist <- unique(left_join(dist, select(lete, recvDeployName, recvLat, recvLon), by = c("recvDeployName2" = "recvDeployName")))
dist <- rename(dist, "recvLat2" = "recvLat", "recvLon2" = "recvLon")
# Get distance (m) between site combos
# get mean lat/lon
dist <- dist %>% group_by(recvDeployName1, recvDeployName2) %>% 
  summarize(recvLat1 = mean(recvLat1),
            recvLon1 = mean(recvLon1),
            recvLat2 = mean(recvLat2),
            recvLon2 = mean(recvLon2)) %>% as.data.frame
dist$distance <- with(dist, latLonDist(recvLat1, recvLon1, recvLat2, recvLon2))
# convert distances from metres to miles
dist$distance <- (dist$distance/1000)/1.609344
write.csv(dist, file = "./data/distances.csv")
```

## Site Visits
To determine the frequency at which Least Terns visited each foraging area, detections at each station were grouped into distinct 'visits'. A visit was categorized as a period of consecutive detections (in minutes) at a single station, or multiple periods of consecutive detections at a single station where there were fewer that 10 minutes between consecutive periods of detection. Detailed visit summaries are available in Appendix \@ref(AppendixC).

```{r hourly, echo = FALSE, message = FALSE, warning = FALSE}
# Get a small data frame (lete.path) of hourly detections for easier data manipulation

# Hourly detections
fun.getpath <- function(df) 
{
  mutate(df, ts.h = as.POSIXct(round(ts, "hours"))) %>%
    group_by(year, motusTagID, runID, ts.h, recvDeployName, recv, 
             tagDeployLon, tagDeployLat, recvLat, recvLon, incubate) %>%
    summarize(max.runLen = max(runLen),
              tot.runLen = sum(runLen)) %>% 
    arrange(motusTagID, ts.h) %>%
    data.frame()
}

lete.path <- fun.getpath(filter(inc, incubate == TRUE))
```

```{r activeReceivers, echo = FALSE, message = FALSE, warning = FALSE, results = "hide"}
#Create data frames of 2017 and 2018 active receivers during tagging period

## list of active receivers
#find active periods for 2017 and 2018
range(filter(lete, year == "2017")$ts)
range(filter(lete, year == "2018")$ts)
# get receiver metadata
tbl.recvDeps <- tbl(sql.motus, "recvDeps")
df.recvDeps <- tbl.recvDeps %>% collect %>% as.data.frame() %>% 
  mutate(tsStart = as_datetime(tsStart, tz = "UTC", 
                               origin = "1970-01-01"), tsEnd = as_datetime(tsEnd, 
                                                                           tz = "UTC", origin = "1970-01-01"))
# for deployments with no end dates, make an end date a year from now
df.recvDeps$tsEnd <- as.POSIXct(ifelse(is.na(df.recvDeps$tsEnd), 
                                       as.POSIXct(format(Sys.time(), "%Y-%m-%d %H:%M:%S")) + 
                                         lubridate::dyears(1), df.recvDeps$tsEnd), tz = "UTC", 
                                origin = "1970-01-01")
# get running intervals for all receiver deployments
siteOp <- with(df.recvDeps, lubridate::interval(tsStart, 
                                                tsEnd))  # get running intervals for each deployment
# set the date range you're interested in
dateRange17 <- lubridate::interval(as.POSIXct("2017-05-15"), 
                                   as.POSIXct("2017-07-04"))
dateRange18 <- lubridate::interval(as.POSIXct("2018-05-08"), 
                                   as.POSIXct("2018-07-10"))
# create new variable 'active' which will be set to 
# TRUE if the receiver was active at some point during the specified date range, 
# and FALSE if not
df.recvDeps17 <- df.recvDeps
df.recvDeps17$active17 <- lubridate::int_overlaps(siteOp, 
                                                dateRange17)
df.recvDeps17 <- select(df.recvDeps17, serno, name, active17, latitude, longitude)
df.recvDeps18 <- df.recvDeps
df.recvDeps18$active18 <- lubridate::int_overlaps(siteOp, 
                                                dateRange18)
df.recvDeps18 <- select(df.recvDeps18, serno, name, active18, latitude, longitude)
# combined data frame showing when each was active
df.recvDepsCombined <- left_join(filter(df.recvDeps17, active17 == TRUE), filter(df.recvDeps18, active18 == TRUE))
df.recvDepsCombined$active18 <- ifelse(is.na(df.recvDepsCombined$active18), "FALSE", df.recvDepsCombined$active18)
df.recvDepsCombined$active <- df.recvDepsCombined$active18
df.recvDepsCombined <- rename(df.recvDepsCombined, "recvDeployName" = "name", "recv" = "serno")
# combine with lete detections for mapping later
tmp <- unique(select(lete.path, recvDeployName, recv, incubate))
tmp <- filter(tmp, incubate == TRUE)
df.recvDepsCombined <- left_join(df.recvDepsCombined, tmp)
```

```{r visits, echo = FALSE, message = FALSE, warning = FALSE}
# Consider birds to be at the nest anytime they are detected with a 
# signal strength below 70 at Exxon Fields (see results), thus foraging will be:
foraging <- filter(inc, !(recvDeployName == "Exxon Fields" & sig <70))
# rounding lat/lons
foraging <- foraging %>% group_by(recvDeployName) %>% 
  mutate(recvLat = mean(recvLat),
            recvLon = mean(recvLon),
            recvLat = mean(recvLat),
            recvLon = mean(recvLon)) %>% as.data.frame

## look at foraging detections
visits <- foraging %>% group_by(motusTagID, mfgID, recvDeployName, recvDeployID, runID, batchID, year) %>%
  summarize(mints = min(ts),
            maxts = max(ts),
            numHits = length(motusTagID))
visits <- as.data.frame(visits)
visits <- visits[with(visits, order(motusTagID, mints)),]
visits <- visits %>%
  mutate(diffsec = as.numeric(mints) - lag(as.numeric(maxts)),
         diffmin = as.numeric(diffsec/60)) %>% as.data.frame
## assign a group number for each group of hits at a station per tag by time
visits <- visits[with(visits, order(motusTagID, mints)),] ## first make sure it's ordered correctly
visits <- visits %>% group_by(motusTagID) %>% mutate(count1 = cumsum(c(1,as.numeric(diff(as.factor(recvDeployName)))!=0))) ## assign a group number each time the time difference from one runID to another is > 10
visits <- visits[with(visits, order(motusTagID, mints)),] ## first make sure it's ordered correctly
visits$diffmin[is.na(visits$diffmin)] <- 0 ## set NA in time differences to 0
visits <- visits %>% group_by(motusTagID, count1) %>% mutate(count2 = c(cumsum(diffmin >10)),
                                                             visitID = paste(count1, count2, sep = ".")) ## now get one value to number each variable

## now get the total time (in minutes) for each "visit"
visits <- visits %>% group_by(motusTagID, mfgID, recvDeployName, recvDeployID, visitID, year) %>%
  summarize(mints = min(mints),
            maxts = max(maxts)) %>% as.data.frame()
visits <- visits %>% mutate(visitLength = as.numeric(difftime(maxts, mints), units = "mins"),
                            date = as.Date(mints),
                            tsRound = as.POSIXct(round(mints, "hours")),
                            visitID = as.numeric(visitID))
visits <- merge(visits, select(df.recvDeps, name, deployID, latitude, longitude, projectID), by.x = "recvDeployID", by.y = "deployID", all.x = TRUE)
## Distance travelled for the visit
visits <- visits[with(visits, order(as.numeric(motusTagID, mints))),]
visits$visit.Dist <- with(visits, latLonDist(lag(latitude), lag(longitude), latitude, longitude)) ## distance for a visit is the distance it took to get there from the last station
# convert distances from metres to miles
visits$visit.Dist <- (visits$visit.Dist/1000)/1.609344
## Sunrise info
visits <- rename(visits, recvDeployLat = latitude, recvDeployLon = longitude, ts = mints)
visits <- rename(visits, recvLat = recvDeployLat, recvLon = recvDeployLon, mints = ts)
visits <- arrange(visits, recvDeployID, motusTagID, mints)
# order site variable based on distance from exxon fields
# first add in distance column
tmp <- filter(dist, recvDeployName1 == "Exxon Fields")
visits <- left_join(visits, select(tmp, recvDeployName2, distance), by = c("recvDeployName" = "recvDeployName2"))
visits <- rename(visits, distFromExxon = distance)
# NOTE: no need to convert distances from metres to miles since 
# they are already in miles from the dist dataframe

visits$recvDeployName <- reorder(visits$recvDeployName, visits$distFromExxon)
  
write.csv(visits, "./data/visits.csv")

# visit length summarized by tagID and receiver
visits.tag <- visits %>% group_by(year, recvDeployName, motusTagID, mfgID) %>%
  summarize(meanLength = mean(visitLength),
            totVisits = length(visitLength)) %>% 
  arrange(year, recvDeployName, motusTagID, totVisits)
write.csv(visits.tag, "./data/visits_tag.csv")

#visit length summarized by receiver
visits.recv <- filter(visits.tag) %>% group_by(year, recvDeployName) %>%
  summarize(meanLength = mean(meanLength),
            meanVisits = mean(totVisits)) %>% 
  arrange(year, recvDeployName)
write.csv(visits.recv, "./data/visits_recv.csv")

```

```{r antBearings, echo = FALSE, message = FALSE, warning = FALSE}
# Create data frame of antenna bearings

# Antenna metadata
tbl.antDeps <- tbl(sql.motus, "antDeps")
df.antDeps <- tbl.antDeps %>% collect %>% as.data.frame()
# merge with receiver metadata
df.antDeps <- left_join(select(df.antDeps, deployID, antennaType, bearing), select(df.recvDeps, deployID, name, latitude, longitude, tsStart, tsEnd), by = "deployID")
# keep only relevant receivers
deps <- unique(df.recvDepsCombined$recvDeployName)
df.antDeps <- filter(df.antDeps, name %in% deps)
df.antDeps <- rename(df.antDeps, "recvDeployName" = "name", "antBearing" = "bearing", "recvDeployID" = "deployID")
```

For each year, an ANOVA test was initially used to determine if the length of visits differed between stations, results indicated that the mean time spent at stations differed significantly between stations for 2017 (p < 0.001) and 2018 (p = 0.001). A Tukey Test was then implemented to discern where the difference lay for both years.
```{r lengthAnova, echo = FALSE, message = FALSE, results = "hide"}
# all years
tmp <- aov(data = filter(visits, recvDeployName %in% c("East Grand Terre", "Exxon Fields", "Grand_Isle", "Wisner")), 
           visitLength ~ recvDeployName)
# 2017
tmp17 <- aov(data = filter(visits, year(date)=="2017"), 
           visitLength ~ recvDeployName)
summary(tmp17)
tt <- TukeyHSD(tmp17, conf.level=0.95) #only Grand_Isle-Exxon is significant
print(tt, digits = 15) # significance of Grand_Isle-Exxon: p = 4.2e-8

#2018
tmp18 <- aov(data = filter(visits, year(date)=="2018"), 
           visitLength ~ recvDeployName)
summary(tmp18)
tt <- TukeyHSD(tmp18, conf.level=0.95) #only Grand_Isle-Exxon is significant
print(tt, digits = 15) # significance of Grand_Isle-Exxon: p = 4.2e-8

#p <- plot(TukeyHSD(tmp, conf.level = 0.95), las=1, col = "red")
```
## Time Partitioning
To determine time partitioning between foraging activities and nest site attendance during the incubation period, individuals that were consistently detected at the nesting site (*Exxon Fields*) were selected. Detection patterns were visually inspected for all tags, and those that were not reliably detected at the nesting site were excluded. In total 5 and 7 tags were selected from 2017 and 2018 respectively. Periods of time during which it could not be confirmed that the bird was incubating were also excluded, for example, the nest of tag 28595 was predated sometime between May 18-21st, and the detection pattern clearly changes during this time, thus detections beyond May 18th were removed for this bird, [Figure 1.1](#fig:siteAttendance). 

To discern behavioural patterns during the incubation period, detections were examined at two levels using the same technique. First, all detections were examined to review general *site* attendance. Second, only detections with a signal strength below 70 were examined as an indication of *nest* attendance, as periods of prolonged low signal strength generally indicate the individual is on the nest. In both cases, only detections at *Exxon Fields* were examined for this subset of tags. 

For both levels, detection periods at *Exxon Fields* were split into discrete 'visits' in the same way site visits were determined; consecutive detections at *Exxon Fields*, or runs of consecutive detections with less than 10 minutes between runs, were classified as one 'visit'. For nest attendance, visits with < 3 detections were also excluded, as we would also expect short periods of decreasing signal strengths as the individual is flying away from the station to forage, as compared to prolonged periods of consistently low signal strength indicative of nest attendance. Furthermore, time of dawn and dusk was used to classify detections as occurring during the day or night.

An ANOVA was used to examine the effect of time of day (day or night) on signal strength and visit length during 2017 and 2018 for site attendance.
```{r, siteAttendance, echo = FALSE, fig.height = 8, fig.cap = "Signal strength vs. time at Exxon Fields for tags with reliable detections during incubation."}
# Create theme for plots
th <- theme_bw() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
                         plot.title = element_text(hjust = 0.5))

# potential tags for incubation detections: 28523, 28526, 28608, 28525, 28520, 23262, 
# 23257, 23248, 23246, 23219, 23218, 28595, 28600
p <- ggplot(filter(inc, motusTagID %in% c(28523, 28526, 28608, 28525, 28520, 23262, 23248, 
                                             23246, 23219, 23218, 28595, 2860), recvDeployName == "Exxon Fields"), aes(ts, sig)) + 
  geom_point() + facet_wrap(~motusTagID, scales = "free") + th + labs(x = NULL, y = "Signal Strength")
# tag 28520, first egg hatched between may 21 and may 24, 
# pattern changes on the 21st so remove detections after the 21st for this bird
# tag 28525 started hatching may 24, so remove detections after this
# tag 28526 failed between may 27-28 so remove detections after 27th
# tag 28595 failed between may 18-21st so remove detections after 18th
nestSite <- inc %>% filter(motusTagID %in% c(28523, 28526, 28608, 28525, 28520, 23262,23257, 23248, 
                                    23246, 23219, 23218, 28595, 28600),
              recvDeployName == "Exxon Fields",
              !(motusTagID == 28520 & ts > as.POSIXct("2018-05-21")),
              !(motusTagID == 28525 & ts > as.POSIXct("2018-05-24")),
              !(motusTagID == 28526 & ts > as.POSIXct("2018-05-27")),
              !(motusTagID == 28595 & ts > as.POSIXct("2018-05-18"))) %>% 
  select(-sunrise, -sunset)
# convert times to CDT since these are Lotek receivers that were recording in local time instead of UTC
nestSite$ts <- (nestSite$ts - 5*60*60)
nestSite$ts <- force_tz(nestSite$ts, tzone = "America/Chicago")
nestSite$date <- as.Date(nestSite$ts, tz = "America/Chicago")

# plot of incubation periods
ggplot(nestSite, aes(ts, sig)) + 
  geom_point(size = 1) + facet_wrap(~motusTagID, scales = "free_x") + th + labs(x = NULL, y = "Signal Strength")

# so with this data, group detections into "visits" to look at mean signal strength
incPeriods <- nestSite %>% group_by(motusTagID, mfgID, recvDeployName, runID, year) %>%
  summarize(mints = min(ts),
            maxts = max(ts),
            numHits = length(motusTagID))
incPeriods <- as.data.frame(incPeriods)
incPeriods <- incPeriods[with(incPeriods, order(motusTagID, mints)),]
incPeriods <- incPeriods %>%
  mutate(diffsec = as.numeric(mints) - lag(as.numeric(maxts)),
         diffmin = as.numeric(diffsec/60)) %>% as.data.frame
incPeriods <- incPeriods[with(incPeriods, order(motusTagID, mints)),] ## first make sure it's ordered correctly
incPeriods$diffmin[is.na(incPeriods$diffmin)] <- 0 ## set NA in time differences to 0
incPeriods <- incPeriods %>% group_by(motusTagID) %>% mutate(visitID = c(cumsum(diffmin >10)), # lump together if time dif is less than 10 mins
                                                             tagVisitID = paste(motusTagID, visitID, sep = ".")) %>% data.frame() ## now get one value to number each variable
## now get the total time (in minutes), and mean signal strength for each "visit"
nestSite <- left_join(nestSite, select(incPeriods, motusTagID, runID, visitID, tagVisitID), by = c("motusTagID", "runID")) 
incPeriods <- nestSite %>% group_by(motusTagID, mfgID, recvDeployName, visitID, tagVisitID, year) %>%
  summarize(mints = min(ts),
            maxts = max(ts),
            numHits = length(motusTagID),
            meanSig = mean(sig),
            minSig = min(sig),
            maxSig = max(sig)) %>% as.data.frame()
incPeriods <- incPeriods %>% mutate(visitLength = as.numeric(difftime(maxts, mints), units = "mins"),
                                    visitID = as.numeric(visitID),
                                    date = as.Date(mints))
# add dawn/dusk times to incPeriods based on first detection for the run
days <- unique(incPeriods$date)
dusk <- getSunlightTimes(date = days, keep = c("dawn", "dusk"),
                         lat = 29.20644, lon = -89.81102, tz = "America/Chicago")
incPeriods <- left_join(incPeriods, dusk, by = "date")
incPeriods$period <- ifelse(incPeriods$mints > incPeriods$dawn & incPeriods$mints < incPeriods$dusk, "day", "night")
# get sunrise etc. info for nestSite dataframe
days <- unique(nestSite$date)
dusk <- getSunlightTimes(date = days, keep = c("sunrise", "sunset", "dawn", "dusk"),
                         lat = 29.20644, lon = -89.81102, tz = "America/Chicago")
nestSite <- left_join(nestSite, dusk, by = "date")
nestSite$period <- ifelse(nestSite$ts > nestSite$dawn & nestSite$ts < nestSite$dusk, "day", "night")

```

```{r partitioningMeans, echo = FALSE, results = "hide"}
## get summaries by day
# longer average length of visit period during night than day in both years (more obvious in 2017)
# mean signal strength was lower during the night as the birds were on the nest
incPeriods %>% group_by(year, period) %>% summarize(meanNumHits = mean(numHits),
                                                   meanLength = mean(visitLength),
                                                   meanSig = mean(meanSig)) %>% as.data.frame()
# Visit Length vs Period ANOVA
tmp <- aov(data = filter(incPeriods, year == 2017), visitLength ~ period) # significant, but incredibly small sample size
tmp <- aov(data = filter(incPeriods, year == 2018), visitLength ~ period) # significant, but incredibly small sample size
summary(tmp) # significant difference 6.6e-11 ***

# Signal Strength vs Period ANOVA
tmp <- aov(data = filter(incPeriods, year == 2017), meanSig ~ period) # significant, but incredibly small sample size
tmp <- aov(data = filter(incPeriods, year == 2018), meanSig ~ period) # significant, but incredibly small sample size
summary(tmp) # significant difference 6.6e-11 ***

# cumulative time spent at station
# on average they spent 198.23 and 144.36 minutes at Exxon in 2017 and 2018 respectively during the day
# on average they spent 251.15 and 133.21 minutes at Exxon in 2017 and 2018 respectively during the night
# to make comparable need to look at how long day and night is (891.71 and 548.28 mins respectively)
# therefore in the day birds were detected at Exxon 22.22% and 16.19% of the time in 2017 and 2018 respectively
# and during the night birds were detected at Exxon 45.81% and 24.30% respectively
tmp <- incPeriods %>% group_by(year, period, date, motusTagID) %>% 
  summarize(totNumHits = sum(numHits),
            totLength = sum(visitLength),
            meanSig = mean(meanSig)) %>% as.data.frame()
tmp %>% group_by(year, period) %>% summarize(meanNumHits = mean(totNumHits),
                                             meanLength = mean(totLength),
                                             meanSig = mean(meanSig)) %>% as.data.frame()

```

```{r nestAttendanceMethods, echo = FALSE, results = "hide"}
# get visit lengths for daylight detections with sig strength under 70
dayInc <- inc %>% filter(motusTagID %in% c(28523, 28526, 28608, 28525, 28520, 23262, 23248, 
                                             23246, 23219, 23218, 28595, 28600),
                           recvDeployName == "Exxon Fields",
                           !(motusTagID == 28520 & ts > as.POSIXct("2018-05-21")),
                           !(motusTagID == 28525 & ts > as.POSIXct("2018-05-24")),
                           !(motusTagID == 28526 & ts > as.POSIXct("2018-05-27")),
                           !(motusTagID == 28595 & ts > as.POSIXct("2018-05-18")),
                           sig < 70) %>% 
  select(-sunrise, -sunset)
# convert times to CDT since these are Lotek receivers that were recording in local time instead of UTC
dayInc$ts <- (dayInc$ts - 5*60*60)
dayInc$ts <- force_tz(dayInc$ts, tzone = "America/Chicago")
dayInc$date <- as.Date(dayInc$ts, tz = "America/Chicago")

# so with this data, group detections into "visits" to look at mean signal strength
dayPeriods <- dayInc %>% group_by(motusTagID, mfgID, recvDeployName, runID, year) %>%
  summarize(mints = min(ts),
            maxts = max(ts),
            numHits = length(motusTagID))
dayPeriods <- as.data.frame(dayPeriods)
dayPeriods <- dayPeriods[with(dayPeriods, order(motusTagID, mints)),]
dayPeriods <- dayPeriods %>%
  mutate(diffsec = as.numeric(mints) - lag(as.numeric(maxts)),
         diffmin = as.numeric(diffsec/60)) %>% as.data.frame
dayPeriods <- dayPeriods[with(dayPeriods, order(motusTagID, mints)),] ## first make sure it's ordered correctly
dayPeriods$diffmin[is.na(dayPeriods$diffmin)] <- 0 ## set NA in time differences to 0
dayPeriods <- dayPeriods %>% group_by(motusTagID) %>% mutate(visitID = c(cumsum(diffmin >10)), # lump together if time dif is less than 10 mins
                                                             tagVisitID = paste(motusTagID, visitID, sep = ".")) %>% data.frame() ## now get one value to number each variable
## now get the total time (in minutes), and mean signal strength for each "visit"
dayInc <- left_join(dayInc, select(dayPeriods, motusTagID, runID, visitID, tagVisitID), by = c("motusTagID", "runID")) 
dayPeriods <- dayInc %>% group_by(motusTagID, mfgID, recvDeployName, visitID, tagVisitID, year) %>%
  summarize(mints = min(ts),
            maxts = max(ts),
            numHits = length(motusTagID),
            meanSig = mean(sig),
            minSig = min(sig),
            maxSig = max(sig)) %>% as.data.frame()
dayPeriods <- dayPeriods %>% mutate(visitLength = as.numeric(difftime(maxts, mints), units = "mins"),
                                    visitID = as.numeric(visitID),
                                    date = as.Date(mints))
# add dawn/dusk times to dayPeriods based on first detection for the run
days <- unique(dayPeriods$date)
dusk <- getSunlightTimes(date = days, keep = c("dawn", "dusk"),
                         lat = 29.20644, lon = -89.81102, tz = "America/Chicago")
dayPeriods <- left_join(dayPeriods, dusk, by = "date")
dayPeriods$period <- ifelse(dayPeriods$mints > dayPeriods$dawn & dayPeriods$mints < dayPeriods$dusk, "day", "night")
# get sunrise etc. info for dayInc dataframe
days <- unique(dayInc$date)
dusk <- getSunlightTimes(date = days, keep = c("sunrise", "sunset", "dawn", "dusk"),
                         lat = 29.20644, lon = -89.81102, tz = "America/Chicago")
dayInc <- left_join(dayInc, dusk, by = "date")
dayInc$period <- ifelse(dayInc$ts > dayInc$dawn & dayInc$ts < dayInc$dusk, "day", "night")

# average length of visit
tmp1 <- filter(dayPeriods, period == "day", numHits > 2) %>% group_by(year) %>% 
  summarize(meanLength = mean(visitLength)) %>% 
  as.data.frame()

# average number of daily visits
tmp <- filter(dayPeriods, period == "day", numHits > 2) %>% group_by(year, date, motusTagID) %>% 
  summarize(totNumHits = sum(numHits),
            totLength = sum(visitLength),
            totVisits = length(motusTagID)) %>% as.data.frame()
tmp2 <- tmp %>% group_by(year) %>% summarize(meanVisits = mean(totVisits)) %>% as.data.frame()
tmp <- left_join(tmp1, tmp2, by = "year")
tmp <- rename(tmp, "Year" = "year", 
              "Mean length of detection periods (mins)" = "meanLength", 
              "Mean number of daily visits" = "meanVisits")

```

