# Appendix - Site Visits {#AppendixC}

```{r tidyr5, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=50), tidy = TRUE)
```
The following tables provide information on site visits to varying degree. [Table C.1](#tab:parameterTableVisits) includes the time and length of each visit for each tag. [Table C.2](#tab:parameterTableVisitsTag) includes the total number and mean length of visits for each tag at each station. [Table C.3](#tab:parameterTableVisitsRecv) includes the total number and mean length of visits by station.

## Unique Visits
The following table shows the length in minutes of each unique visit:
```{r parameterTableVisits, echo = FALSE, out.extra='angle=90'}
param.table <- dplyr::select(read.csv("./data/visits.csv", stringsAsFactors=FALSE), 
                             motusTagID, mfgID, recvDeployName, year, mints, maxts, 
                             visitLength, recvLat, recvLon, distFromExxon)
param.table <- param.table %>% 
  mutate(Location = paste(round(recvLat, digits = 2), round(recvLon, digits = 2), sep = ", "),
         visitLength = round(visitLength, digits = 2))
param.table <- rename(param.table, "Motus Tag ID" = motusTagID, "Lotek ID" = mfgID,
                      Receiver = recvDeployName, Year = year, 
                      "First Detection" = mints,
                      "Final Detection" = maxts, "Visit Length (mins)" = visitLength, 
                      "Distance from Exxon Fields (m)" = distFromExxon)
#kable(select(param.table, "Motus Tag ID", "Lotek ID", Receiver, Location, 
#                    "Distance from Exxon Fields (m)", "First Detection", "Final Detection", 
#                    "Visit Length (mins)"), 
#             booktabs = TRUE, longtable = TRUE,
#             caption = 'Time and length of each unique visit for each tag')
pander(select(param.table, "Motus Tag ID", "Lotek ID", Receiver, Location,
              "Distance from Exxon Fields (m)", "First Detection", "Final Detection","Visit Length (mins)"), 
      booktabs = TRUE, caption = '(\\#tab:parameterTableVisits) Time and length of each unique visit for each tag', 
      split.cell = 20, split.table = Inf, justify = 'left')
```


## Visits by tag and station
For each year and tag, the following table shows the total number and mean length (mins) of visits at each station:
```{r parameterTableVisitsTag, echo = FALSE, out.extra='angle=90'}
param.table <- dplyr::select(read.csv("./data/visits_tag.csv", stringsAsFactors=FALSE), 2:7, 10)
param.table <- param.table %>% mutate(meanLength = round(meanLength, digits = 2),
                                      seVisits = round(seVisits), digits = 2)
param.table <- arrange(param.table, motusTagID, recvDeployName)
param.table <- rename(param.table,  "Motus Tag ID" = motusTagID, "Lotek ID" = mfgID,
                      Receiver = recvDeployName, Year = year, 
                      "Mean Length of Visits (mins)" = meanLength, 
                      "Total Number of Visits" = totVisits,
                      "Standard Error Visits" = seVisits)
#kable(select(param.table, Year, "Motus Tag ID", "Lotek ID", Receiver, 
#             "Total Number of Visits", "Mean Length of Visits (mins)"), 
#      booktabs = TRUE, caption = 'Total number and mean length of visits (mins) for each tag by receiver', longtable = TRUE)
pander(select(param.table, Year, "Motus Tag ID", "Lotek ID", Receiver, 
             "Total Number of Visits", "Mean Length of Visits (mins)", "Standard Error Visits"), 
      booktabs = TRUE, caption = '(\\#tab:parameterTableVisitsTag) Total number and mean length of visits (mins) for each tag by receiver', split.cell = 20, split.table = Inf, justify = 'left')
```


## Visits by Station
For each year and receiver, the following table shows the mean number and length (mins) of visits of all tags.

```{r parameterTableVisitsRecv, echo = FALSE, out.extra='angle=90'}
param.table <- dplyr::select(read.csv("./data/visits_recv.csv", stringsAsFactors=FALSE), 2:4,7,8,11)
param.table <- param.table %>% mutate(meanlength = round(meanlength, digits = 2),
                                      seLength = round(seLength, digits = 2),
                                      meanVisits = round(meanVisits, digits = 2),
                                      seVisits = round(seVisits), digits = 2)
param.table <- rename(param.table, Year = year, Receiver = recvDeployName, 
                      "Mean Length of Visits (mins)" = meanlength, 
                      "Standard Error Length of Visits" = seLength,
                      "Mean Number of Visits" = meanVisits,
                      "Standard Error Number of Visits" = seVisits)
#kable(param.table, 
#      booktabs = TRUE, caption = 'Mean length (mins) and number of visits of all tags at each receiver by year', longtable = TRUE) 
pander(param.table, 
      booktabs = TRUE, caption = '(\\#tab:parameterTableVisitsRecv) Mean length (mins) and number of visits of all tags at each receiver by year', split.cell = 30, split.table = Inf, justify = 'left') 
```
